Causal Compatibility Inequalities Admitting Quantum Violations in the Triangle Structure

### 2018-08-01 CCIAQVTS/CCIAQVTS.tex
Made changes to the figures in order to have the figure files match their appearance in the document.

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure}
    \author{Thomas C. Fraser}
    \email{tfraser@perimeterinstitute.ca}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    \affiliation{University of Waterloo, Waterloo, Ontario, Canada, N2L 3G1}
    \author{Elie Wolfe}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    % \email{ewolfe@perimeterinstitute.ca}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are incompatible with particular assumption about classical causal structure. Given a causal structure of unknown classicality, the presence of such correlations certifies the non-classical nature of the causal structure in a device independent fashion. In structures where all parties share a common resource, these non-classical correlations are also known as non-local correlations. Any constraint satisfied by all correlations which \textit{are} classically compatible with a given causal structure defines a causal compatibility criterion. Such criteria were recently derived for the Triangle structure [\href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}] in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical Triangle structure is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution in this work, however, is the derivation of new causal compatibility inequalities for the Triangle structure which \textit{do} admit quantum violation. This provides the first robust-to-noise witness of quantum correlations in the Triangle structure. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    % \tableofcontents
    % \setlength{\parskip}{1em}

@@

    If a given distribution $\prob[\nodes]$ defined over \textit{all} of the nodes $\nodes$ of $\graph$ satisfies \cref{eq:compatibility_as_factorization}, then $\prob[\nodes]$ is said to be \term{compatible with $\graph$}. The conditional distributions in \cref{eq:compatibility_as_factorization} (i.e. $\{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes\}$) are referred to as a set of \term{causal parameters for $\graph$}. If a distribution $\prob[\nodes]$ cannot be factorized according to \cref{eq:compatibility_as_factorization}, $\prob[\nodes]$ is said to be \term{incompatible} with $\graph$.

    When one is specified with a joint distribution $\prob[\nodes]$ defined over all nodes of a causal structure $\graph$, it is possible to completely determine whether or not $\prob[\nodes]$ is compatible with $\graph$ by computing the causal parameters induced by $\prob[\nodes]$ and checking the equality of \cref{eq:compatibility_as_factorization}.

    A challenge, however, is presented when one is supplied with a partial observation, i.e. a joint distribution $\prob[\nodes_{O}]$ where $\nodes_{O} \subset \nodes$ is some subset of variables referred to as the \term{observable nodes} $\nodes_O$. In such cases, $\prob[\nodes_{O}]$ does not induce a unique set of causal parameters for $\graph$ and \cref{eq:compatibility_as_factorization} can not be verified by direction calculation. Instead, compatibility between $\graph$ and $\prob[\nodes_{O}]$ depends on the existence or non-existence of a set of causal parameters for $\graph$ such that $\prob[\nodes_{O}] = \prod_{n \in \nodes_{O}} \prob[n \mid \Pa[\graph]{n}]$. The complementary, unobservable nodes are termed \term{latent nodes} $\nodes_L = \nodes \setminus \nodes_{O}$, and should be understood as hidden random variables that are unknowable either by some fundamental process or cannot be measured due to other limitations.

@@

            \includegraphics[trim={0 2.2cm 0 0}, clip, scale=0.6]{figure_fritz_dist_plotted_bits.pdf}
            % \vspace{-0.2in}
            \[ \probplotvalue{253, 253, 84} = \f{1}{32}\br{2 - \sqrt{2}} \qquad \probplotvalue{11, 1, 0} = \f{1}{32}\br{2 + \sqrt{2}} \qquad \probplotvalue{255, 255, 255} = 0\]

@@

    \begin{figure}
    \begin{nscenter}
            \includegraphics[scale=0.6,trim={0 2.2cm 0 0.4in},clip]{figure_noisy_yet_non_local_fritz.pdf}
            % \vspace{-0.2in}
            \[ \probplotvalue{253, 253, 84} = 0.018 \qquad \probplotvalue{11, 1, 0} = 0.099 \qquad \probplotvalue{255, 255, 243} = 0.0013 \]
            \caption{$\s{N}_{0.085}$: A noisy yet still non-classical variant of the Fritz distribution.}
            \label{fig:noisy_fritz}
    \end{nscenter}
    \end{figure}

@@

    Although there are numerous techniques that can used when parameterizing quantum states and measurements~\cite{Petz_2015, Hedemann_2013,Fujii_2005,James_2001,Grasmair_2014,Neilsen_Chaung_2011}, a single technique by \citet{Spengler_2010_Unitary} was found to be most computationally suitable for our purposes. \Citet{Spengler_2010_Unitary} demonstrated that all $d\times d$ unitary matrices $U$ can be parameterized without degeneracy as follows:
    \[ U = \bs{\prod_{m=1}^{d-1} \br{\prod_{n=m+1}^{d} \exp\br{i P_n \lambda_{n,m}}\exp\br{i \si_{m,n} \lambda_{m,n}}}} \cdot \bs{\prod_{l=1}^{d} \exp\br{iP_l \lambda_{l,l}}}  \eq \label{eq:spengler_unitary} \]
    Where the real valued parameters $\la = \bc{\la_{n,m} \mid n,m \in 1, \ldots, d}$ have periodicities $\la_{m,n} \in \bs{0, \f{\pi}{2}}$ for $m < n$ and $\la_{m,n} \in \bs{0, 2 \pi}$ for $m \geq n$. Moreover, $P_l$ are one-dimensional projective operators $P_l = \ket{l}\bra{l}$ and the $\si_{m,n}$ are generalized anti-symmetric $\si$-matrices $\sigma_{m,n} = -i \ket{m}\bra{n} +i \ket{n}\bra{m}$ where $1 \leq m < n \leq d$. This parameterization has the useful feature that each of the real-valued parameters $\la_{n,m}$ a direct and intuitive physical affect on each element of a computational basis $\bc{\ket{1}, \ldots, \ket{d}}$. Explicitly, $\exp\br{i \si_{m,n} \lambda_{m,n}}$ applies a rotation to the sub-space spanned by $\ket{m}$ and $\ket{n}$ for $m < n$. Analogously, $\exp\br{i P_n \lambda_{n,m}}$ generates the relative phase between $\ket{m}$ and $\ket{n}$ for $m > n$ and $\exp\br{iP_l \lambda_{l,l}}$ fixes the global phase of $\ket{l}$. Finally, although not explicitly mentioned in~\cite{Spengler_2010_Unitary}, it is possible to remove the reliance on the computationally expensive matrix exponential operations~\cite{Moler_2003} in \cref{eq:spengler_unitary} and replace them with elementary trigonometric functions in the parameters $\la_{m,n}$.

@@

    \begin{figure}
    \begin{nscenter}
        \includegraphics[scale=0.6,trim={0 0 0 0.4in},clip]{figure_plotted_dist_I_1_max_violation.pdf}
        \caption{A quantum probability distribution of the Triangle structure that maximizes violation of $I\tsb{Web}$. Notice that this distribution has precisely the same possibilistic structure as the Fritz distribution, as might be expected, since $I\tsb{Web}$ was specifically generated to witness the non-classicality of $\prob[\fritz]$.}
        \label{fig:maximum_violation_I_1}
    \end{nscenter}
    \end{figure}
    \begin{figure}
    \begin{nscenter}
        \includegraphics[scale=0.6,trim={0 0 0 0.4in},clip]{figure_plotted_dist_I_3_max_violation.pdf}
        \caption{A quantum probability distribution of the Triangle structure that maximizes violation of $I\tsb{SymmetricWeb}$. Notice that this distribution has precisely the same (asymmetric!) possibilistic structure as the Fritz distribution, even though $I\tsb{SymmetricWeb}$ itself is symmetric with respect to permutations of the variables $A, B, C$. }
        \label{fig:maximum_violation_I_3}
    \end{nscenter}
    \end{figure}

>>>>

\begin{document}
    \title{Causal Compatibility Inequalities Admitting Quantum Violations in the Triangle Structure}
    \author{Thomas C. Fraser}
    \email{tfraser@perimeterinstitute.ca}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    \affiliation{University of Waterloo, Waterloo, Ontario, Canada, N2L 3G1}
    \author{Elie Wolfe}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    % \email{ewolfe@perimeterinstitute.ca}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are incompatible with particular assumption about classical causal structure. Given a causal structure of unknown classicality, the presence of such correlations certifies the non-classical nature of the causal structure in a device independent fashion. In structures where all parties share a common resource, these non-classical correlations are also known as non-local correlations. Any constraint satisfied by all correlations which \textit{are} classically compatible with a given causal structure defines a causal compatibility criterion. Such criteria were recently derived for the Triangle structure (E. Wolfe \textit{et al.}, \href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}) in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical Triangle structure is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution in this work, however, is the derivation of causal compatibility inequalities for the Triangle structure which do admit quantum violation. This provides a robust-to-noise witness of quantum correlations in the Triangle structure. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    % \tableofcontents
    % \setlength{\parskip}{1em}

@@

    If a given distribution $\prob[\nodes]$ defined over \textit{all} of the nodes $\nodes$ of $\graph$ satisfies \cref{eq:compatibility_as_factorization}, then $\prob[\nodes]$ is said to be \term{compatible with $\graph$}. The conditional distributions in \cref{eq:compatibility_as_factorization} (i.e. $\{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes\}$) are referred to as a set of \term{causal parameters for $\graph$}. If a distribution $\prob[\nodes]$ cannot be factorized according to \cref{eq:compatibility_as_factorization}, $\prob[\nodes]$ is said to be \term{incompatible} with $\graph$. When one is specified with a joint distribution $\prob[\nodes]$ defined over all nodes of a causal structure $\graph$, it is possible to completely determine whether or not $\prob[\nodes]$ is compatible with $\graph$ by computing the causal parameters induced by $\prob[\nodes]$ and checking the equality of \cref{eq:compatibility_as_factorization}. A challenge, however, is presented when one is supplied with a partial observation, i.e. a joint distribution $\prob[\nodes_{O}]$ where $\nodes_{O} \subset \nodes$ is some subset of variables referred to as the \term{observable nodes} $\nodes_O$. In such cases, $\prob[\nodes_{O}]$ does not induce a unique set of causal parameters for $\graph$ and \cref{eq:compatibility_as_factorization} can not be verified by direction calculation. Instead, compatibility between $\graph$ and $\prob[\nodes_{O}]$ depends on the existence or non-existence of a set of causal parameters for $\graph$ such that $\prob[\nodes_{O}] = \prod_{n \in \nodes_{O}} \prob[n \mid \Pa[\graph]{n}]$. The complementary, unobservable nodes are termed \term{latent nodes} $\nodes_L = \nodes \setminus \nodes_{O}$, and should be understood as hidden random variables that are unknowable either by some fundamental process or cannot be measured due to other limitations.

@@

            \includegraphics{figure_fritz_dist_plotted_bits.pdf}

@@

    \begin{figure}
    \begin{nscenter}
            \includegraphics{figure_noisy_yet_non_local_fritz.pdf}
            \caption{$\s{N}_{0.085}$: A noisy yet still non-classical variant of the Fritz distribution.}
            \label{fig:noisy_fritz}
    \end{nscenter}
    \end{figure}

@@

    Although there are numerous techniques that can used when parameterizing quantum states and measurements~\cite{Petz_2015, Hedemann_2013,Fujii_2005,James_2001,Grasmair_2014,Neilsen_Chaung_2011}, a single technique by \citet{Spengler_2010_Unitary} was found to be most computationally suitable for our purposes. \Citet{Spengler_2010_Unitary} demonstrated that all $d\times d$ unitary matrices $U$ can be parameterized without degeneracy as follows:
    \[ U = \bs{\prod_{m=1}^{d-1} \br{\prod_{n=m+1}^{d} \exp\br{i P_n \lambda_{n,m}}\exp\br{i \si_{m,n} \lambda_{m,n}}}} \times \bs{\prod_{l=1}^{d} \exp\br{iP_l \lambda_{l,l}}}  \eq \label{eq:spengler_unitary} \]
    Where the real valued parameters $\la = \bc{\la_{n,m} \mid n,m \in 1, \ldots, d}$ have periodicities $\la_{m,n} \in \bs{0, \f{\pi}{2}}$ for $m < n$ and $\la_{m,n} \in \bs{0, 2 \pi}$ for $m \geq n$. Moreover, $P_l$ are one-dimensional projective operators $P_l = \ket{l}\bra{l}$ and the $\si_{m,n}$ are generalized anti-symmetric $\si$-matrices $\sigma_{m,n} = -i \ket{m}\bra{n} +i \ket{n}\bra{m}$ where $1 \leq m < n \leq d$. This parameterization has the useful feature that each of the real-valued parameters $\la_{n,m}$ a direct and intuitive physical affect on each element of a computational basis $\bc{\ket{1}, \ldots, \ket{d}}$. Explicitly, $\exp\br{i \si_{m,n} \lambda_{m,n}}$ applies a rotation to the sub-space spanned by $\ket{m}$ and $\ket{n}$ for $m < n$. Analogously, $\exp\br{i P_n \lambda_{n,m}}$ generates the relative phase between $\ket{m}$ and $\ket{n}$ for $m > n$ and $\exp\br{iP_l \lambda_{l,l}}$ fixes the global phase of $\ket{l}$. Finally, although not explicitly mentioned in~\cite{Spengler_2010_Unitary}, it is possible to remove the reliance on the computationally expensive matrix exponential operations~\cite{Moler_2003} in \cref{eq:spengler_unitary} and replace them with elementary trigonometric functions in the parameters $\la_{m,n}$.

@@

    \begin{figure}
    \begin{nscenter}
        \includegraphics{figure_plotted_dist_I_1_max_violation.pdf}
        \caption{A quantum probability distribution of the Triangle structure that maximizes violation of $I\tsb{Web}$. Notice that this distribution has precisely the same possibilistic structure as the Fritz distribution, as might be expected, since $I\tsb{Web}$ was specifically generated to witness the non-classicality of $\prob[\fritz]$.}
        \label{fig:maximum_violation_I_1}
    \end{nscenter}
    \end{figure}
    \begin{figure}
    \begin{nscenter}
        \includegraphics{figure_plotted_dist_I_3_max_violation.pdf}
        \caption{A quantum probability distribution of the Triangle structure that maximizes violation of $I\tsb{SymmetricWeb}$. Notice that this distribution has precisely the same (asymmetric!) possibilistic structure as the Fritz distribution, even though $I\tsb{SymmetricWeb}$ itself is symmetric with respect to permutations of the variables $A, B, C$. }
        \label{fig:maximum_violation_I_3}
    \end{nscenter}
    \end{figure}
